Carryover effects in free recall reveal how past experiences influence memories of future experiences

### 2023-01-16 paper/main.tex
backing up draft of context effects discussion section

@@

@@

% connections to prior work: 
%     - context effects: 
%     - priming: 
%     - situation models:

@@

@@

>>>>

\subsection*{Recap: visual feature manipulations}

@@

\subsection*{Recap: order manipulations}

@@

% connections to prior work: 
%     - context effects: role of irrelevant features, temporal context and order effects
%     - priming: being "primed" to 
%     - situation models:

\subsection*{Context effects on memory performance and organization}

In everyday experience, each moment's unique blend of contextual features
(where we are, who we are with, what else we are thinking of at the time, what
else we experience nearby in time, etc.) plays an important role in how we
interpret, experience, and remember that moment, and how we relate it to our
other experiences~\citep[e.g., for review see ][]{Mann20}. What are the
analogues of real-world contexts in laboratory tasks like the free recall
paradigm employed in our study? In general, modern formal accounts of free
recall~\citep{Kaha20} describe context as comprising a mix of (a) features
pertaining to or associated with each item and (b) other items and thoughts
experienced nearby in time, e.g., that might still be ``lingering'' in the
participant's thoughts at the time they study the item. Item features can
include semantic properties (i.e., features related to the item's meaning),
lexicographic properties (i.e., features related to the item's letters),
sensory properties (i.e., feature related to the item's appearance, sound,
smell, etc.), emotional proprties (i.e., features related to how meaningful the
item is, whether the item evokes positive or negative feelings, etc.),
utility-related properties (e.g., features that describe how an item might be
used or incorporated into a particular task or situation), and more.
Essentially any aspect of the participant's experience that can be
characterized, measured, or otherwise described can be considered to influence
the participant's mental context at the moment they experience that item.
Temporally proximal features include aspects of the participant's internal or
external experience that are \textit{not} specifically occuring at the moment
they encounter an item, but that nonetheless influence how they process the
item. Thoughts related to percepts, goals, expectations, other experiences, and
so on that might have been cued (directly or indirectly) by the participant's
recent experiences prior to the current moment all fall into this category.
Internally driven mental states, such as thinking about an experience unrelated
to the experiment, also fall into this category.

Contextual features need not be intentionally or consciously perceived by the
participant to affect memory, nor do they need to be relevant to the task
instructions or the participant's goals. Incidental factors such as font
color~\citep{JonePyc14}, background color~\citep{IsarIsar07}, inter-stimulus
images~\citep{GersEtal13, MannEtal16, ChiuEtal21}, background
sounds~\citep{BeamJone10, SahaSmit14}, secondary tasks~\citep{MasiSaha14,
PolyEtal09}, and more can all impact how the participant remembers, and
organizes in memory, lists of studied items.

Consistent with this prior work, we found that participants are sensitive to
task-irrelevant visual features. We also found that changing the dynamics of
those task-irrelevant visual features (in the reduced (early) and reduced
(late) conditions) \textit{also} affects participants' memories. This suggests
that it is not only the contextual features themselves that affect memory, but
also the \textit{dymamics} of context-- i.e., how the contextual features
associated with each item change over time.

\subsection*{Priming effects on memory performance and organization}

\subsection*{Expectation, event boundaries, and situation models}

@@

\subsection*{Theoretical implications}

\subsection*{Potential applications}

@@

\subsection*{Concluding remarks}
